---
title: "Tidyverse Example with Rock Dataset"
author: "Amish Rasheed"
date: "2024-11-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

In this example, we’ll use the Tidyverse functions `group_by` and `summarize` to analyze the built-in rock dataset in R. This dataset includes measurements of rock samples, such as area, perimeter, shape, and permeability. We'll group the data by a categorical representation of the shape variable and calculate summary statistics.

# Load Tidyverse

```{r}
library(tidyverse)
```

# Load the built-in rock dataset

```{r}
data("rock")
```

## The rock dataset contains four columns:

**area**: Area of the rock sample in square pixels.
**peri**: Perimeter of the rock sample in pixels. 
**shape**: Shape of the rock sample. 
**perm**: Permeability of the rock sample.

# Example: Summarize by a rounded 'shape' value

```{r}
rock_summary <- rock %>%
  mutate(shape_category = round(shape, 1)) %>%  # Round to one decimal place
  group_by(shape_category) %>%                 
  summarize(
    avg_area = mean(area),                   
    avg_perm = mean(perm),                   
    count = n()                              
  )

# Plot the summarized data
ggplot(rock_summary, aes(x = factor(shape_category), y = avg_area)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(
    title = "Average Area by Shape Category",
    x = "Shape Category (Rounded to 1 Decimal Place)",
    y = "Average Area"
  ) +
  theme_minimal()
```

# Conclusion
In this example, I demonstrated how to use group_by and summarize to compute summary statistics for grouped data using the Tidyverse package. I also created a bar plot to visualize the average area across different shape categories.
