---
title: "BW_Gather"
author: "Ben Wolin"
date: "2024-11-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Demonstrating the Gather() functionality in tidyr
Gather takes multiple columns and collapses into key-value pairs, duplicating all other columns as needed. You use gather() when you notice that you have columns that are not variables.

## Pulling in an example from the word bank data catalog.
https://energydata.info/dataset/world-climate-watch/resource/1631a4e8-a59a-4026-aa36-162df9b15340

```{r}
library(tidyr)
Consumption_raw <- read.csv('https://raw.githubusercontent.com/bwolin99/TestRepo/refs/heads/main/606%20Final/Meat_Consumption.csv',row.names = NULL)
head(Consumption_raw)
```
As we can see in this dataset there are many columns indicating the year in which the measurement was taken. We would like to group these years into one column and the total emissions into the other.

## Grouping with Gather()
```{r}
Consumption <- gather(Consumption_raw, key = 'year', value = 'Meat_Consumption',2:28)
head(Consumption)
```

Now we turned 28 columns into 3 using the Gather function.
